---
title: "Data processing: Interpolating bite scar size distribution"
author: "Hannah Rempel"
date: "Last updated 10/January/2023"
output: html_document
---

```{r Setup}
if (!require(doParallel)) install.packages("doParallel")
if (!require(here)) install.packages("here") 
if (!require(purrr)) install.packages("purrr") 
if (!require(tidyverse)) install.packages("tidyverse") 

packages <- c("doParallel", "here", "purrr", "tidyverse")
sapply(packages, require, character.only = T)
```

```{r Reading in data}
scar_df_bites <- read_csv(here("data/regional_coral_scar_data.csv")) %>% 
  filter(percent_new>0) #filtering to only include colonies w/ recent predation scars
```

In the field, we recorded the mininum, median, and maximum size of scars, as well as the total abundance of scars. For colonies with scars, the following chunk interpolates that size of scars based these values.

For colonies with <5 scars:   
*   If there is 1 scar, it is assigned the value of scar_med_mm2
*   If there are 2 scars, 1 is assigned the value of scar_min_mm2 and 1 scar_max_mm2,
*   If there are 3 scars, 1 is assigned the value scar_min_mm2, 1 scar_med_mm2, 1 scar_max_mm2
*   If there are 4 scars, 1 is assigned the value scar_min_mm2, 2 scar_med_mm2, 1 scar_max_mm2

For colonies with 5+ scars:   
*   Min and Max: 1 is assigned the value of scar_min_mm2 and 1 scar_max_mm2
*   Median: An abundance of x=[n_scars]/2 are assigned the value of scar_med_mm2
*   Lower mid: An abundance of x=(([n_scars]/2)-2)/2 are assigned the the average of scar_min_mm2 and scar_med_mm2, that is: (scar_min_mm2 + scar_med_mm2)/2
*   Upper mid: An abundance of x=(([n_scars]/2)-2)/2 are assigned the the average of scar_max_mm2 and scar_med_mm2, that is: (scar_max_mm2 + scar_med_mm2)/2

Note that not all scar abundances from this interpolated distribution are whole numbers, they can also be in quarters (0.25, 0.5, 0.75, 1.25, etc). These interpolated abundances are denoted in the column 'num' below. For example, for a colony with 5 scars, 1 is assigned the min and max area, 2.5 the median area, and 0.25 the upper and lower mid values. This results in a total count of 5 scars, but within a given scar area, not all values are in whole numbers (the smallest fraction is a quarter). 
```{r Interpolating scar sizes}
#creates vector to of l (scar area) and a (scar abundance) numbered 1-5:
scar_area_mm_sqabund <- c('area_min', 'num_min', 'area_lower_mid', 'num_lower_mid', 'area_med', 'num_med', 'area_upper_mid','num_upper_mid', 'area_max', 'num_max')

#adds scar_area_mm_sqgth and abundance variables to scar_df_bites
scar_df_bites[scar_area_mm_sqabund] <- NA

#triangle distribution to create abundances at scar_area_mm_sqgths
for (i in 1:dim(scar_df_bites)[1]) {
  
   if(scar_df_bites$n_scars[i] == 1) {
    scar_df_bites$area_med[i] <- scar_df_bites$scar_med_mm2[i]; scar_df_bites$num_med[i] <- 1;

    } else if(scar_df_bites$n_scars[i] == 2) {
    scar_df_bites$area_min[i] <- scar_df_bites$scar_min_mm2[i]; scar_df_bites$num_min[i] <- 1;
    scar_df_bites$area_max[i] <- scar_df_bites$scar_max_mm2[i]; scar_df_bites$num_max[i] <- 1 
    
    } else if(scar_df_bites$n_scars[i] == 3)  {
      scar_df_bites$area_min[i] <- scar_df_bites$scar_min_mm2[i]; scar_df_bites$num_min[i] <- 1;
      scar_df_bites$area_med[i] <- scar_df_bites$scar_med_mm2[i]; scar_df_bites$num_med[i] <- 1;
      scar_df_bites$area_max[i] <- scar_df_bites$scar_max_mm2[i]; scar_df_bites$num_max[i] <- 1
      
      } else if(scar_df_bites$n_scars[i] == 4)  {
        scar_df_bites$area_min[i] <- scar_df_bites$scar_min_mm2[i]; scar_df_bites$num_min[i] <- 1;
        scar_df_bites$area_med[i] <- scar_df_bites$scar_med_mm2[i]; scar_df_bites$num_med[i] <- 2;
        scar_df_bites$area_max[i] <- scar_df_bites$scar_max_mm2[i]; scar_df_bites$num_max[i] <- 1  
        
        } else if(scar_df_bites$n_scars[i] >= 5)  {
          
          #one scar assigned minimum area
          scar_df_bites$area_min[i] <- scar_df_bites$scar_min_mm2[i]; 
          scar_df_bites$num_min[i] <- 1;
          
          scar_df_bites$area_lower_mid[i] <- round((scar_df_bites$scar_min_mm2[i]+scar_df_bites$scar_med_mm2[i])/2); 
          scar_df_bites$num_lower_mid[i] <- (scar_df_bites$n_scars[i]/2 - 2)/2;
          
          scar_df_bites$area_med[i] <- scar_df_bites$scar_med_mm2[i]; 
          scar_df_bites$num_med[i] <- scar_df_bites$n_scars[i]/2;
          
          scar_df_bites$area_upper_mid[i] <- round((scar_df_bites$scar_max_mm2[i]+scar_df_bites$scar_med_mm2[i])/2);
          scar_df_bites$num_upper_mid[i] <- scar_df_bites$num_lower_mid[i]; #same abundance of scars assigned to upper mid as the lower mid
          
          #one scar assigned maximum area
          scar_df_bites$area_max[i] <- scar_df_bites$scar_max_mm2[i]; 
          scar_df_bites$num_max[i] <- 1
  }
}

scar_df_bites1 <- scar_df_bites
scar_df_bites1$scar_area_mm_sq <- scar_df_bites$area_min
scar_df_bites1$num <- scar_df_bites$num_min

scar_df_bites2 <-scar_df_bites
scar_df_bites2$scar_area_mm_sq <- scar_df_bites$area_lower_mid
scar_df_bites2$num <- scar_df_bites$num_lower_mid

scar_df_bites3 <-scar_df_bites
scar_df_bites3$scar_area_mm_sq <- scar_df_bites$area_med
scar_df_bites3$num <- scar_df_bites$num_med

scar_df_bites4 <-scar_df_bites
scar_df_bites4$scar_area_mm_sq <- scar_df_bites$area_upper_mid
scar_df_bites4$num <- scar_df_bites$num_upper_mid

scar_df_bites5 <-scar_df_bites
scar_df_bites5$scar_area_mm_sq <- scar_df_bites$area_max
scar_df_bites5$num <- scar_df_bites$num_max

interpolated_scar_data <- rbind(scar_df_bites1, scar_df_bites2, scar_df_bites3, scar_df_bites4, scar_df_bites5) %>%
  #filtering out instances where there wasn't a scar of a particular size
  filter(!is.na(scar_area_mm_sq)) %>% 
  dplyr::select(region, site, date, transect, colony_id, coral_species, l_cm, w_cm, percent_live, n_scars, percent_new, scar_area_mm_sq, num)%>% 
  #calculating coral area based on SA of a hemisphere* percent living coral tissue
  mutate(col_area_cm_sq = ((pi * (l_cm/2) * (w_cm/2))) * (percent_live/100),
         scar_area_cm_sq = scar_area_mm_sq/100) %>%
  dplyr::select(-scar_area_mm_sq) 
```

The following chunks generate bootstrapped estimates of the distribution of the standing stock of new bite scars from the dataframe containing scar size distribution and percent new scars.

Reason: For each coral, we estimated % new versus old scars, but not for each indivual scar. Our model of OANN tissue regneration is based the area of new scars, so to apply our model to predict tissue loss from the standing stock of scars we first need to use bootstrapping to randomly sample new scars from the dataset based on the % new scars per colony and repeat this process 1000 times to generate estimates of the distribution of new scars.

This will result in a dataframe where n_rows(predicted_tissue_loss_df) == n_iterations. Each row will be predicted tissue loss from randomly assigning scars on a coral colony to be new versus old, and calculating tissue loss for the new scars. Since this is done, for instance, a thousand times, we get reasonable estimates of actual tissue loss like we would if we knew which individual scars were actually new or old.

This code in run in parallel to reduce the time it takes to run 

```{r Bootstrapped standing stock of scars}
set.seed(4561)

#preparing data for bootstrapped sampling
n_new_df <- interpolated_scar_data %>% 
  mutate(num_rnd=num*4) %>% #multiplying by four so data is in whole numbers
  uncount(num_rnd) %>% #converting to long form
  mutate(n_new_scars = ceiling((percent_new/100)* n_scars)) %>% #creating variable for number of new scars
  dplyr::select(region, site, transect, coral_species, colony_id, l_cm, w_cm, percent_live, scar_area_cm_sq, n_new_scars) %>%
  arrange(colony_id)

#index of colony ids and n scars to sample used in bootstrappig
index_of_cols <- n_new_df %>%
  select(colony_id, n_new_scars) %>% 
  distinct() %>%
  arrange(colony_id)
#empty df of summarized_predicted_tissue_loss_df for outer loop
new_scar_standing_stock <- tibble(
  region = character(),
  site = character(),
  transect = integer(),
  coral_species = character(),
  colony_id = integer(),
  l_cm = numeric(),
  w_cm = numeric(),
  percent_live = numeric(),
  scar_area_cm_sq = numeric(),
  n_new_scars= numeric()
  )

#### Function to samples scars ####

standing_stock_sample <- function(n) {  
  
  for (iteration in 1:n){
    est_new_scar_stock_df <-n_new_df %>%
    group_split(colony_id) %>%
    map2_dfr(index_of_cols$n_new_scars, ~ slice_sample(.x, n = .y, replace=FALSE)) %>%
    mutate(iteration_n=iteration)
    
    new_scar_standing_stock <- dplyr::bind_rows(est_new_scar_stock_df, new_scar_standing_stock) 
    }
    return(new_scar_standing_stock)
}  

#### Sampling scar standing stock 10000 times using parallel processing ####

no_cores <- detectCores() - 1  #creates a number of cores on your computer -1
doParallel::registerDoParallel(cores=no_cores) #call so that code will run on your no_cores -1
cl <- makeCluster(no_cores, type="FORK")
bootstrapped_scar_df <- parLapply(cl, 10000, standing_stock_sample)[[1]]  #applying function in parallel
stopCluster(cl)
```


```{r writing bootstrapped data to csv}
bootstrapped_scar_df %>%  
  dplyr::reframe(max(iteration_n)) #quick check

write.csv(bootstrapped_scar_df, here("data/bootstrapped_new_scar_data.csv"), 
          fileEncoding = "UTF-8", 
          row.names=FALSE)
```
