---
title: "Estimating mean corallivory rates by parrotfish species from published studies"
author: "Hannah Rempel"
date: "2023-03-02"
output: html_document
---

```{r Setup}
if (!require(here)) install.packages("here") 
if (!require(janitor)) install.packages("janitor") 
if (!require(tidyverse)) install.packages("tidyverse") 

packages <- c("here", "janitor", "tidyverse")
sapply(packages, require, character.only = T)
```

```{r Reading in STX and FLA data}
fla_grazing_raw <- read_csv(here("supplements/supp_data/published_fish_foraging_rates_raw/FLA_foraging_data.csv")) %>%
  janitor::clean_names()

stx_grazing_raw <- read_csv(here("supplements/supp_data/published_fish_foraging_rates_raw/STX_foraging_data.csv")) %>%
  janitor::clean_names()
```

All Florida observations from the Adam et al. dataset were of initial phase (IP) individuals, as described in Adam et al. 2015 and Burkepile et al. 2019
```{r Cleaning and summarizing parrotfish behavioral data from Florida}
fla_grazing_raw %>% 
  distinct(food_category)

fla_coral_predation <- fla_grazing_raw %>% 
  filter(!food_category %in% c("Not_observed", "Unknown")) %>%
  filter(food_category=="Coral") %>%
  group_by(location, species, fish_id) %>%
  summarise(n_bites_coral=sum(bites))

fla_corallivory_by_fish <- fla_grazing_raw %>% 
  group_by(location, species, fish_id) %>%
  filter(!food_category %in% c("Not_observed", "Unknown")) %>%
  summarise(n_bites=sum(bites)) %>% 
  left_join(fla_coral_predation, by = join_by(location, species, fish_id)) %>%
  mutate(n_bites_coral=replace_na(n_bites_coral, 0),
         pct_bites_coral=n_bites_coral/n_bites*100)

fla_corallivory_by_species <- fla_corallivory_by_fish %>%
  mutate(region="Florida", phase="IP") %>%
  group_by(region, species, phase) %>%
  summarise(mean_corallivory=mean(pct_bites_coral),
            n_observed=n()) %>%
  arrange(-mean_corallivory)
```

```{r Cleaning and summarizing parrotfish behavioral data from St. Croix}
stx_grazing_raw %>% 
  distinct(food_specific)

stx_coral_predation <- stx_grazing_raw %>% 
  filter(! food_specific %in% c("unknown", "out of sight")) %>%
  filter(food_specific %in% c("porites porites","porites astreoides", "orbicella annularis")) %>%
  group_by(location, species, phase, date, time_start) %>% #to summarize observations by each unique fish
  summarise(n_bites_coral=sum(bites)) %>%
  mutate(species=str_replace(species, "sparisoma", "Sparisoma"),
         species=str_replace(species, "scarus", "Scarus"))

stx_corallivory_by_fish <- stx_grazing_raw %>% 
  group_by(location, species, phase, date, time_start) %>%
  filter(! food_specific %in% c("unknown", "out of sight")) %>%
  summarise(n_bites=sum(bites)) %>% 
  left_join(stx_coral_predation, by = join_by(location, species, phase, date, time_start)) %>%
  mutate(n_bites_coral=replace_na(n_bites_coral, 0),
         pct_bites_coral=n_bites_coral/n_bites*100) %>%
  mutate(species=str_replace(species, "sparisoma", "Sparisoma"),
         species=str_replace(species, "scarus", "Scarus"))

stx_corallivory_by_species  <- stx_corallivory_by_fish  %>%
  mutate(region="St. Croix") %>%
  group_by(region, species, phase) %>%
  summarise(mean_corallivory=mean(pct_bites_coral),
            n_observed=n()) %>%
  arrange(-mean_corallivory) 
```

Writing summarized Florida and St. Croix parrotfish corallivory rate data to csv file
```{r Merging FLA and STX data and writing to CSV}
fla_stx_corallivory_by_species <- fla_corallivory_by_species %>% 
  bind_rows(stx_corallivory_by_species)

write_csv(fla_stx_corallivory_by_species, here("supplements/supp_data/STX_FLA_summarized_corallivory_rates.csv"))
```

Mean percent of bites on coral from the Cardoso et al. 2009 (Fig. 2) and sample sizes per species (Fig. 1) were extracted using WebPlotDigitizer and added into a combined excel sheet with the summarized data from Florida and St. Croix. These data are read in below.

```{r Reading in combined summaries of parrotfish corallivory rates}
corallivory_rates_combined <- read_csv(here("supplements/supp_data/corallivory_by_parrotfish_sp.csv"))
```

Calculated the mean corallivory rates across the datasets weighted by the sample size for each dataset.
```{r Calculating the mean rates of corallivory by parrotfish species}
mean_corallivory_by_pf_sp <- corallivory_rates_combined %>%
  group_by(species) %>%
  uncount(replicates) %>% #uncounts observations by # of replicates to calculate mean weighted by the sample size
  summarise(mean_pct_bites_on_coral=mean(as.numeric(pct_bites_on_coral)),
            replicates=n()) %>%
  arrange(-mean_pct_bites_on_coral) 
```

```{r Saving Table S2 to csv}
write_csv(mean_corallivory_by_pf_sp, here("supplements/table_s3_published_parrotfish_corallivory_rates.csv"))
```